---
title: "Assignment 2: Coding Basics"
author: "Charles :-)"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. 

#make call to seq function, display to console
seq_1_to_100 <- seq(1,100,4)
seq_1_to_100

#2. 

#call mean and median functions, store in variables, display to console
seq_mean <- mean(seq_1_to_100)
seq_median <- median(seq_1_to_100)
seq_mean
seq_median

#3.

#use an ifelse() to check if stored mean is greater than median
#print statemnts accordingly to console
ifelse(seq_mean > seq_median, "Mean is greater", "Median is greater or equal to the mean")

```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}

#String Vector of names
names <- c("Bob", "Tom", "Jack", "Jill", "Jasmine")

#number vector of scores
scores <- c(87, 69, 50, 32, 45)

#boolean vector with associated pass/fail 
pass_fail <- c(TRUE, TRUE, TRUE, FALSE, FALSE)

#generate data frame
df = data.frame(names, scores, pass_fail)

```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: In a data frame, you can have different kinds of data, but in a matrix all elements must be the same data type.

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement.

11. Apply your function to the vector with test scores that you created in number 5.

```{r}

check_pass <- function(scores_input){
  for (score in scores_input) {
    ifelse(score>=50, print(TRUE), print(FALSE))
  }
}

check_pass(scores)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: I used the ifelse() function nested in a for loop. For each score in the inputted vector of scores the ifelse() statement compared the score with 50. If the score was greater than or equal to the passing grade of 50 then TRUE was printed. Otherwise, FALSE was printed. I decided to use the ifelse() function rather than if{} else{} brackets for cleaner more concise code.
